
```{r}
library(sf)
library(tidyverse)
library(tmap)
library(spdep)

```


```{r}
res2021 <- st_read('data/results/res2021.gpkg')
res2022 <- st_read('data/results/res2022.gpkg')
res2023 <- st_read('data/results/res2023.gpkg')


```
```{r}

# 确保 res2021 是 sf 对象
res2021 <- st_make_valid(res2021)
res2022 <- st_make_valid(res2022)
res2023 <- st_make_valid(res2023)

tm_shape(res2021) +
    tm_polygons("casualty_people_rate",
        style="fisher",
        palette="PuOr",
        midpoint=NA,
        title="casualty_people_rate")

```

neighbours list
```{r}
library(spdep)

coordsW2021 <- res2021%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW2021,axes=TRUE)

#create a neighbours list
res2021_nb <- res2021 %>%
  poly2nb(., queen=T)
res2022_nb <- res2022 %>%
  poly2nb(., queen=T)
res2023_nb <- res2023 %>%
  poly2nb(., queen=T)

#plot them
plot(res2021_nb, st_geometry(coordsW2021), col="red")
#add a map underneath
plot(res2021$geom, add=T)

```

weights matrix
```{r}
#create a spatial weights matrix from these weights
res2021.lw <- res2021_nb %>%
  nb2listw(., style="C")
res2022.lw <- res2022_nb %>%
  nb2listw(., style="C")
res2023.lw <- res2023_nb %>%
  nb2listw(., style="C")

```

Moran’s I
Moran’s I test tells us whether we have clustered values (close to 1) or dispersed values (close to -1),
```{r}
I_res2021_Global_people_Density <- res2021 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  moran.test(., res2021.lw)
I_res2021_Global_length_Density <- res2021 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  moran.test(., res2021.lw)

I_res2022_Global_people_Density <- res2022 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  moran.test(., res2022.lw)
I_res2022_Global_length_Density <- res2022 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  moran.test(., res2022.lw)

I_res2023_Global_people_Density <- res2023 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  moran.test(., res2023.lw)
I_res2023_Global_length_Density <- res2023 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  moran.test(., res2023.lw)

I_res2021_Global_people_Density
I_res2021_Global_length_Density
I_res2022_Global_people_Density
I_res2022_Global_length_Density
I_res2023_Global_people_Density
I_res2023_Global_length_Density
```

Geary’s C 
Geary’s C tells us whether similar values or dissimilar values are clustering
Moran's I：

Moran's I 的统计量更容易理解，并且具有较好的理论基础，特别是在理论研究中，广泛用于测试空间数据的自相关性。
Geary's C：
Geary's C 侧重于局部差异，理论上较为复杂。其结果可能会受到离群值和局部特征的影响，解释起来有时较为困难。

```{r}
# C_LWard_Global_people_Density <- 
#   res2021 %>%
#   pull(casualty_people_rate) %>%
#   as.vector()%>%
#   geary.test(., res2021.lw)
# 
# C_LWard_Global_length_Density <- 
#   res2021 %>%
#   pull(casualty_length_rate) %>%
#   as.vector()%>%
#   geary.test(., res2021.lw)
# 
# C_LWard_Global_people_Density
# C_LWard_Global_length_Density

```

 Local Moran’s I
```{r}
I_res2021_local_people_Density <- res2021 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  localmoran(., res2021.lw) %>% 
  as_tibble()

I_res2021_local_length_Density <- res2021 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  localmoran(., res2021.lw) %>% 
  as_tibble()

#what does the output (the localMoran object) look like?
slice_head(I_res2021_local_people_Density, n=5)
```
```{r}
res2021 <- res2021 %>%
  mutate(casualty_people_rate_I =
           as.numeric(I_res2021_local_people_Density$Ii))%>%
  mutate(casualty_people_rate_Iz=
           as.numeric(I_res2021_local_people_Density$Z.Ii))%>%
  mutate(casualty_length_rate_I =as.numeric(I_res2021_local_length_Density$Ii))%>%
  mutate(casualty_length_rate_Iz =as.numeric(I_res2021_local_length_Density$Z.Ii))
```
 

mapping
```{r}
library(RColorBrewer)
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

MoranColours<- rev(brewer.pal(8, "RdGy"))

tm_shape(res2021) +
    tm_polygons("casualty_people_rate_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, casualty_people_rate")

tm_shape(res2021) +
    tm_polygons("casualty_length_rate_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, casualty_length_rate")
```

Local Getis Ord G

```{r}
Gi_res2021_Local_people_Density <- res2021 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  localG(., res2021.lw)

Gi_res2021_Local_length_Density <- res2021 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  localG(., res2021.lw)

head(Gi_res2021_Local_people_Density)
head(Gi_res2021_Local_length_Density)

res2021 <- res2021 %>%
  mutate(casualty_people_rate_G = as.numeric(Gi_res2021_Local_people_Density)) %>%   mutate(casualty_length_rate_G = as.numeric(Gi_res2021_Local_length_Density))

```
mapping
```{r}
library(RColorBrewer)

GIColours<- rev(brewer.pal(8, "RdBu"))

#now plot on an interactive map
tm_shape(res2021) +
    tm_polygons("casualty_people_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Gi*, casualty_people_rate")
```
```{r}

#now plot on an interactive map
tmap_mode('plot')

tm_shape(res2021) +
    tm_polygons("casualty_length_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G")
```
Local Moran's I：
评估每个空间单元（如区域）与其邻居之间的空间自相关性，揭示 局部空间聚集模式。
正的局部 Moran's I 值：表明该区域及其邻居具有相似的属性值（如高-高聚集或低-低聚集）。
负的局部 Moran's I 值：表明该区域与邻居的属性值存在显著差异（如高-低或低-高）。

Local Getis-Ord G：确定空间数据中的热点和冷点，即显著高值或低值的集群，而不考虑值之间的正负相关性。
高G值：表明该区域是一个 热点（高值聚集区）。
低G值：表明该区域是一个 冷点（低值聚集区）。
因此就用Getis-Ord G值：

```{r}
Gi_res2021_Local_people_Density <- res2021 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  localG(., res2021.lw)

Gi_res2021_Local_length_Density <- res2021 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  localG(., res2021.lw)

head(Gi_res2021_Local_people_Density)
head(Gi_res2021_Local_length_Density)

res2021 <- res2021 %>%
  mutate(casualty_people_rate_G = as.numeric(Gi_res2021_Local_people_Density)) %>%   mutate(casualty_length_rate_G = as.numeric(Gi_res2021_Local_length_Density))

Gi_res2022_Local_people_Density <- res2022 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  localG(., res2022.lw)

Gi_res2022_Local_length_Density <- res2022 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  localG(., res2022.lw)

res2022 <- res2022 %>%
  mutate(casualty_people_rate_G = as.numeric(Gi_res2022_Local_people_Density)) %>%   mutate(casualty_length_rate_G = as.numeric(Gi_res2022_Local_length_Density))


Gi_res2023_Local_people_Density <- res2023 %>%
  pull(casualty_people_rate) %>%
  as.vector()%>%
  localG(., res2023.lw)

Gi_res2023_Local_length_Density <- res2023 %>%
  pull(casualty_length_rate) %>%
  as.vector()%>%
  localG(., res2023.lw)

res2023 <- res2023 %>%
  mutate(casualty_people_rate_G = as.numeric(Gi_res2023_Local_people_Density)) %>%   mutate(casualty_length_rate_G = as.numeric(Gi_res2023_Local_length_Density))

```


mapping
```{r}
library(RColorBrewer)

breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
GIColours<- rev(brewer.pal(8, "RdBu"))

#now plot on an interactive map
t1 <- tm_shape(res2021) +
    tm_polygons("casualty_people_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Rate (2021)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))
t1
tmap_save(t1, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Rate (2021).png')
```

```{r}
#now plot on an interactive map
tmap_mode('plot')

t2 <- tm_shape(res2021) +
    tm_polygons("casualty_length_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Location Rate (2021)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))

tmap_save(t2, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Length Rate (2021).png')
t2
```

```{r}
t3 <- tm_shape(res2022) +
    tm_polygons("casualty_people_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Rate (2022)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))

tmap_save(t3, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Rate (2022).png')
t3
```

```{r}
t4 <- tm_shape(res2022) +
    tm_polygons("casualty_length_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Location Rate (2022)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))

tmap_save(t4, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Length Rate (2022).png')
t4

```

```{r}
t5 <- tm_shape(res2023) +
    tm_polygons("casualty_people_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Rate  (2023)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))

tmap_save(t5, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Rate (2023).png')
t5
```

```{r}
t6 <- tm_shape(res2023) +
    tm_polygons("casualty_length_rate_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Local Getis-Ord G") +
  tm_compass(position = c("right", "top"),type = "arrow") + 
  tm_scale_bar(position = c("center", "bottom")) +
  tm_layout(title = "Hot and cold spots of Casualty Location Rate (2023)", 
            legend.position = c("left", "bottom"),
            title.position = c("center", "top"))

tmap_save(t6, 'hot_cold_spots_graphs/Hot and cold spots of Casualty Length Rate (2023).png')
t6
```

